{\helveticaitalic{Supplementary Material}}:\\ \helvetica{ Benchmarking neuromorphic systems with Nengo}

### 2015-07-01 paper/paper.tex
Full edit by Trevor
Filled in most sections now (except stuff at the end).
Ready to rerun benchmarks and regenerate plots.

\usepackage{url,lineno}
\linenumbers

@@

  Abstract

  \tiny
  % 5--8 keywords
  \keyFont{ \section{Keywords:}  }
\end{abstract}

@@

???expand Other software that do similar things:
many neural simulators exist, and have tests,
but don't have the high-level API
that Nengo has.
On the other hand, some simulators
have high-level APIs
(Topographica?) but a) aren't functional
and b) can't use diverse backends.

??? In the subsequent sections,
we detail the architecture of Nengo's
frontend and backend,
and describe what any backend
is required to implement.
We explain Nengo's testing framework,
including descriptions of the testing fixtures
used to collect and visualize benchmarks.
We then list the metrics that are collected
in this paper,
and the backends on which we collect metrics.
Finally, we show and discuss
the results of collecting those metrics
for each backend.

??? Name backends and give brief intro to them
in intro. Can then refer to them in other sections.

@@

There are two key features of Nengo 2.0
that enable rapid benchmarking of neuromorphic systems.
The first is a decoupling of model creation and simulation,
resulting in a platform independent frontend interface
to any backend that implements a certain set of requirements.
The second is a ``fuzzy'' test suite ensuring that
Nengo can be used to create models that solve cognitive tasks.
This test suite makes ample use of test fixtures
to enable data collection while testing on arbitrary backends.

@@

Nengo has a strict separation between
the frontend and backend.
The frontend exposes a modelling interface
that uses Python to define models concisely.
Backends are responsible for transforming
those frontend objects into code that
can be run on the target platform.
While each backend must be exposed
through Python, this requirement
does not significantly limit
flexibility in the backend.
A backend can be implemented in C and exposed
through Python bindings,
or can be run as a separate process
managed by the Python backend,
with data transmitted to Python through sockets
or other inter-process communication protocols.

@@

Nengo contains five frontend objects
that validate and store symbolic information
about the neural model to be simulated.
All neural models defined with Nengo
are built with these five building blocks,
including Spaun, a model that performs
eight cognitive tasks with 2.5 million neurons
that use visual input to produce motor output
\cite{spaun}.
A backend only needs to be aware of these
five objects to run Nengo models,
and therefore run the benchmarks discussed in this paper.

@@

The \textit{Probe} denotes that a particular
quantity in the simulation should be recorded
at a particular rate.
The object to be probed must be specified;
the attribute of that object,
a sampling rate, and a synapse model for filtering
can optionally be specified.
The \textit{Probe} provides the main mechanism
for data collection during run time.

@@

Finally, the \textit{Network} is a
container for the other four objects.
The only parameter on the network
is an optional integer seed;
setting this should make
all of the random factors
in the model deterministic.
Aside from managing the random seed,
the network is responsible for maintaining
a network-specific set of default parameters
for the four other objects.
This results in much shorter and less error-prone
model creation scripts.

An example script showing how Nengo can
express a functional network
with ??? neurons and ??? connections
(representing ??? connection weights)
in ??? lines of code can be seen
in Figure~\ref{???}.

\subsubsection{Backend requirements}

@@

These three methods and one attribute
make up the public interface
that Nengo tests expect from backends.
While the existence of the methods
and attribute are necessary
for every Nengo model,
they are not sufficient;
backends are not required
to implement all of the neuron models
or advanced features that
are available in all other backends.
Backends are responsible for informing
the user when model cannot be implemented
on that backend.
The compliance metric (see section ???)
explicitly tracks the details of what
types of models can be implemented
by each backend.
The details of how each backend
implements high-level frontend model
are out of the scope of this paper.

@@

As of version 2.0.1, Nengo's full test suite contains 618 tests.
Many tests are unit tests
that ensure the frontend API operates as we expect.
212 tests construct a simulator instance,
run the simulator, and test the output of that model
(as exposed through \texttt{trange}
and \texttt{data})
and therefore can be considered ``functional'' tests.
For the remainder of this paper,
we will focus only on these 212 functional tests.

@@

Nengo's test suite employs the \texttt{pytest}
testing framework \cite{???}.
\texttt{pytest} enables
expressive testing of pluggable components
(such as backends)
through what are called ``test fixtures.''
Fixtures are exposed to the test suite
as arguments that can be provided
to any test function.
\texttt{pytest} inspects the function signature,
and passes an appropriate value
to the function.
This allows for boilerplate code
to be specified once,
and run for any test that might use it.
The example script in Figure~\ref{???}
demonstrates how these fixtures
look in test scripts.

@@

\paragraph{\texttt{Simulator}}
allows for tests to be run with multiple backends.
As detailed in section ???,
each backend must expose a simulator class
that accepts a network.
Every test uses this test fixture rather than
an actual class so that
the backend-specific simulator class is used
when the test is run.
Each backend's test suite loads
the tests defined in Nengo,
but replaces the implementation of
this fixture with their own fixture
returning that particular backend's simulator class.

\paragraph{\texttt{plt}, \texttt{logger}, \& \texttt{analytics}}
allow tests to save artifacts from test runs.
These artifacts are how accuracy and speed metrics
are collected, and how the figures
in this paper are generated.
The \texttt{plt} fixture exposes the Matplotlib \cite{???}
\texttt{pyplot} interface to a test function.
These functions can then analyze and plot
the data generated in a simulation
(exposed from a simulator's \texttt{data} attribute)
to visualize the activity in a network.
Figures are saved in a specified or backend-specific directory.
The \texttt{logger} fixture exposes a logging interface
to save arbitrary text to a specified
or backend-specific directory;
these are useful for summary statistics
that can inspected manually.
For large amounts of data that
requires more analysis,
\texttt{analytics} fixture exposes an interface
to save arbitrary data (in the form of NumPy arrays)
to specified or backend-specific directories.

\paragraph{\texttt{analytics\_data}}
enables comparative testing between
two runs of the same test.
The two runs might represent two versions
of the same backend---one before a
speedup and one after---or two
entirely different backends.
The \texttt{analytics\_data} fixture
provides the results from those two
test runs to a single test function,
allowing for comparisons
between the two results saved
with the \texttt{analytics} fixture.
See Figure~\ref{???} for a
concrete example of using the
\texttt{analytics\_data} fixture.

@@

\subsection{Metrics collected}

\paragraph{Compliance}
is the number of tests that a particular backend
passes successfully, relative to the reference
backend, which must pass all tests.
While this metric can give an indication
of how many features a particular backend supports,
it does not take into account how commonly used
a particular feature is;
for this reason, a backend that has relatively
low compliance can still be very useful
in many situations.

@@

@@

Unlike the compliance metric,
accuracy and speed are only collected
for tests that define an accuracy metric
(i.e., those that use the \texttt{analytics} fixture).
In this study, we focus on four such tests
that address central functioning aspects
of a wide variety of large-scale briain models
(see \cite{NEF or SPA book or both?}).
??? Each test has several parameters
that can be varied
in order to benchmark extreme cases
(e.g., representing many dimensions
with few neurons).
We test each model with parameters
typically used in large-scale models,
but make these parameterized models
available at \url{???github url}
for those who want to explore extreme cases.

@@

??? include code for this example, as an appendix?

??? Also include appendix with version numbers used?

\paragraph{2. Two-dimensional product.}
In this model,
the scalar product is computed
from a two-dimensional ensemble
comprised of 100 LIF neurons.
Since the product is nonlinear,
we use a space-filling curve
(the Hilbert curve \cite{???})
as the input signal to ensure that
we sample the entire two-dimensional space.
This model tests each backend's
ability to compute a nonlinear function,
albeit a low-dimensional one.
Despite the low dimensionality,
two-dimensional products are frequently used
in large-scale models;
it is the primary operation needed to support
symbol-like processing using neuron-like elements,
for example \cite{spaun or SPA book}.

@@

\paragraph{4. Basal Ganglia sequence.}
In this model, brain structures
known collectively as the basal ganglia
are contructed from 4900 LIF neurons,
organized such that they iterate
through a repeating set of actions.
This model has been used to investigate
action selection and learning \cite{stewart, bekolay}
and the switching time between actions
has been closely mapped to human decision making
\cite{stewart 2010}.
This benchmark tests the ability
to construct this model,
and evaluates the time needed
to transition between actions.

@@

\paragraph{1. Reference (\texttt{nengo}).}
The reference backend is designed to run quickly
on any general purpose computer
by using NumPy \cite{???}
for fast vectorized computations.
It is included with the \texttt{nengo} frontend
as \texttt{nengo.Simulator}.
When appropriate, other backends will be benchmarked
relative to the reference backend,
which has the most features but does not target
specialized hardware.

@@

\paragraph{3. OpenCL (\texttt{nengo\_ocl}).}
The OpenCL backend uses the Open Computing Language
(OpenCL; \cite{???})
to run Nengo models on many different computing devices,
including graphical processing units (GPUs),
and field-programmable gate arrays (FPGAs).
In contrast to the distilled backend,
it is designed to run Nengo models
as quickly as possible,
using fast general purpose computing devices
like GPUs and any optimizations available,
at the cost of code readability.

\paragraph{4. Brainstorm software (\texttt{nengo\_brainstorm}).}
The Brainstorm backend is a software implementation
of a new neuromorphic chip
based partly on Neurogrid \cite{???}
currently in development
by the Brain in Silicon lab
at Stanford University.
The software backend does not aim for speed;
instead, it attempts to emulate the hardware
in order to test its applicability
to large-scale neural models.
If the emulated hardware can perform well,
then it follows that actual hardware
will also perform well,
but will be much faster.

\paragraph{5. SpiNNaker hardware (\texttt{nengo\_spinnaker}).}
The SpiNNaker backend \cite{Mundy et al., 2015 in press}
targets the eponymous neuromorphic hardware
developed by Furber et al. \cite{10.1109/JPROC.2014.2304638}.
In contrast to the Brainstorm backend,
this backend targets physical neuromorphic hardware,
and therefore is concerned
both with accuracy and speed.

\subsection{Hardware used for running benchmarks}

??? ask Xuan

SpiNN-5 board

@@

Figure~\ref{fig:02} gives accuracy results
for the two-dimensional product model.
In this case, while all of the backends
perform well, the SpiNNaker backend
performs significantly worse than the
other four backends.
Its performance, however, is still well within
acceptable ranges for the two-dimensional product model.

Figure~\ref{fig:03} gives accuracy results
for the controlled oscillator model.
The reference and OpenCL backends perform well here,
as do the distilled and Brainstorm backends.
The SpiNNaker backend performs well in most cases,
but some instances of the model
are significantly less accurate
than the majority of instances.

Figure~\ref{fig:04} gives accuracy results
for the basal ganglia sequence model.
Again, all of the backends perform similarly,
except for the SpiNNaker backend.
All other backends have a transition time
around 43 milliseconds, which matches
the results from Stewart et al. 2010 \cite{???}
for simple actions.
The SpiNNaker backend has a transition time
around 55 milliseconds, which is slower
than the time reported by Stewart et al.
for simple actions, but faster than the
time reported for complex actions.

@@

Figures~\ref{fig:05} and \ref{fig:06} show
the build and run speeds for all backends, respectively.
While build time is not critical to optimize,
it is worth noting that the reference backend
has consistently fast builds.
The OpenCL backend is also usually fast,
but can be slow for certain types of networks
(e.g., the chained communication channel model).
The distilled and Brainstorm backends---which
have a very similar build process---become
slow when dealing with moderately sized models
(e.g., the basal ganglia sequence model,
which contains 4900 neurons).
The SpiNNaker backend incurs an unavoidable cost
in setting up the hardware,
though it builds the largest model faster than
the distilled and Brainstorm backends.

The run time results are of the utmost importance
when evaluating backends for different applications,
especially those that require real-time interaction,
such as robotics. In general, all backends
perform adequately for the three smaller models.
The OpenCL backend is significantly slower for small
models, like the product and controlled oscillator models.
SpiNNaker, on the other hand, is slower than other backends
for these models, but only because its goal
is to run in real-time even if it is possible to run faster.
(??? confirmed: always runs at real time)
The most important result is for the largest model,
the basal ganglia sequence.
In this model, the distilled and Brainstorm backends
performed very poorly
(though neither has speed as a primary goal).
The reference backend also did not perform well,
operating at nearly eight times slower than real-time.
The OpenCL backend performed well, operating at around
four times slower than real-time.
The best performance was seen by the SpiNNaker backend,
however, which operated at around 1.5 times real-time
on this moderately-sized model.

??? For SpiNNaker, place and route is most costly part of
preparing a sim, not necessarily interfacing with hardware

??? Add something about overhead; calling ``run'' still
results in writing data to the board. Actual simulation
occurs in real-time for SpiNNaker; can safely interact
with robotics, neuromorphic hardware, etc

Despite the fast run times for
the OpenCL and SpiNNaker backends,
we also investigated the possibility
that probes were slowing these backend down,
since probes require moving data from
their native platform (GPU or SpiNNaker board)
to the CPU. However, we found that
removing probes from these models
made almost no impact on their run time
(see Figure~\ref{Fig:07}).

@@

The most important finding from these benchmarks
is that specialized hardware (including GPUs)
can result in significantly faster run times
for certain models,
with little to no cost in accuracy.
The conditions under which the OpenCL backend
performs slower than the reference backend
(e.g., the product and controlled oscillator models)
bear closer inspection to determine the source
of the slowdown.

Similarly, we should investigate
the model instances in which the SpiNNaker backend
is less accurate (e.g., in the controlled oscillator model)
or different from other backends
(e.g., in the basal ganglia sequence model).
??? Product difference might be due to different saturation
properties in SpiNNaker ???

Above and beyond the benchmarking results,
however, we believe that the primary contribution
of this study is to provide evidence that Nengo
has a tested, stable, productive frontend
that can target multiple backends,
and therefore provides a platform
for benchmarking neuromorphic backends,
and other neural simulators.
While this is similar to the goal
of other projects, such a PyNN \cite{???},
Nengo is unique in focusing on large-scale simulations,
rather than attempting to support
detailed single-neuron models
(though this capability is still possible in Nengo;
see \cite{???}).

@@

???concluding remarks

@@

>>>>

\usepackage{url,lineno}
\linenumbers
\usepackage{inconsolata}
\usepackage{listings}
\lstset{language=Python}
\lstset{basicstyle=\ttfamily,breaklines=true}

@@

  Nengo is a software package for designing and simulating
  large-scale neural models.
  Nengo is architected such that the same Nengo model
  can be simulated on any of several
  Nengo backends with few to no modifications.
  Backends translate a model to specific platforms,
  which include GPUs and neuromorphic hardware.
  Nengo also contains a large test suite
  that can be run with any backend
  and focuses primarily on functional performance.
  We propose that Nengo's large test suite
  can be used to benchmark neuromorphic hardware's
  functional performance and speed simulating
  in an efficient, unbiased, and future-proof manner.
  We implement four benchmark models and show that
  Nengo can collect several benchmark metrics
  that identify situations in which
  some backends perform differently
  or significantly faster than others.

  \tiny
  % 5--8 keywords
  \keyFont{ \section{Keywords:} Nengo, benchmarking, neuromorphic hardware }
\end{abstract}

@@

Several Nengo backends exist or are in active development.
In this paper, we will run benchmarks
on five backends.
The reference, distilled, and Brainstorm backends
run on general purpose computers,
and focus on aspects other than speed.
The OpenCL backend aims to be a fast backend
that still works on general purpose computers,
but can also take advantage of
graphical processing units (GPUs).
The SpiNNaker backend uses
SpiNNaker neuromorphic hardware to
simulate models in real time
for natural interaction with
physical sensors and actuators.

In the subsequent sections,
we detail the architecture of Nengo's
frontend and backend,
and describe what any backend
is required to implement.
We describe Nengo's testing framework,
including explanations of the testing fixtures
used to collect and visualize benchmarks.
We then list the metrics
that are collected in this paper,
and give further details on the backends benchmarked.
Finally, we show and discuss
the results of collecting those metrics
for each backend.

@@

There are two key features of Nengo
(starting from version 2.0)
that enable rapid benchmarking of neuromorphic systems.
The first is a decoupling of model creation and simulation,
resulting in a platform independent frontend interface
to any backend that implements a certain set of requirements.
The second is a ``fuzzy'' test suite ensuring that
Nengo can be used to create models that solve cognitive tasks.
The test suite makes ample use of test fixtures
to enable data collection while testing on arbitrary backends.

@@

Nengo has a strict separation between
frontend and backend.
The frontend exposes a modelling interface
that uses Python to define models concisely.
Backends are responsible for transforming
those frontend objects into code that
can be run on the target platform.
While each backend must be exposed
through Python, this requirement
does not significantly limit
flexibility in the backend.
A backend can be implemented in C and exposed
through Python bindings,
or can be run as a separate process
managed by the Python backend,
with data transmitted to Python through sockets
or other inter-process communication protocols.

@@

Nengo contains five frontend objects
that validate and store symbolic information
about the neural model to be simulated.
All neural models defined with Nengo
are built with these five building blocks,
including Spaun, a model that performs
eight cognitive tasks with 2.5 million neurons
that use visual input to produce motor output
\cite{eliasmith2012}.
A backend only needs to be aware of these
five objects to run Nengo models,
and therefore run the benchmarks discussed in this paper.

@@

The \textit{Probe} provides the main mechanism
for data collection during run time
by denoting that a particular
quantity in the simulation should be recorded
at a particular rate.
The object to be probed must be specified;
the attribute of that object,
a sampling rate, and a synapse model for filtering
can optionally be specified.

@@

Finally, the \textit{Network} is a
container for the other four objects.
The only parameter on the network
is an optional integer seed;
setting this should make
all of the random factors
in the model deterministic.
Aside from managing the random seed,
the network is responsible for maintaining
a network-specific set of default parameters
for the four other objects.
This results in shorter and less error-prone
model creation scripts.

An example script showing how Nengo can
express a functional network
with 500 neurons and 6 connections
(representing 50,000 connection weights
and 100 direct current injection sites)
in under 20 lines of code
can be seen in Appendix~\ref{App:01}.

\subsubsection{Backend requirements}
\label{Sec:Backend}

@@

These three methods and one attribute
make up the public interface
that Nengo tests expect from backends.
While the existence of these methods
and attribute are necessary
for every Nengo model,
backends are not required
to implement all of the neuron models
or advanced features that
are available in all other backends.
Backends are responsible for informing
the user when a model cannot be implemented
on that backend.
The compliance metric (see Section~\ref{Sec:Metrics})
explicitly tracks what
types of models can be implemented
by each backend.
The details of how each backend
implements high-level frontend models
are out of the scope of this paper.

@@

As of version 2.0.1, Nengo's full test suite contains 618 tests.
Many tests are unit tests
that ensure the frontend API operates as expected.
212 tests construct a simulator instance,
run the simulator, and test the output of that model
(as exposed through \texttt{trange}
and \texttt{data})
and therefore can be considered ``functional'' tests.
For the remainder of this paper,
we will focus only on these 212 functional tests.

@@

Nengo's test suite employs the \texttt{pytest}
testing framework \footnote{Available at \url{http://pytest.org/}.}.
\texttt{pytest} enables
expressive testing of pluggable components
(such as backends)
through what are called ``test fixtures.''
Fixtures are exposed to the test suite
as arguments that can be provided
to any test function.
\texttt{pytest} inspects the function signature,
and passes an appropriate value
to the function.
This allows for boilerplate code
to be specified once,
and run for any test that might use it.
The example script in Appendix~\ref{App:01}
demonstrates how these fixtures
look in test scripts.

@@

\paragraph{\texttt{Simulator}}
allows for tests to be run with multiple backends.
As detailed in Section~\ref{Sec:Backend},
each backend must expose a simulator class
that accepts a network.
Every test uses this test fixture rather than
an actual class so that
the backend-specific simulator class is used
when the test is run.
Each backend's test suite loads
the tests defined in Nengo,
but replaces the implementation of
this fixture with their own fixture
returning that particular backend's simulator class.

\paragraph{\texttt{plt}, \texttt{logger}, \& \texttt{analytics}}
allow tests to save artifacts from test runs.
These artifacts are how accuracy and speed metrics
are collected, and how the figures
in this paper are generated.
The \texttt{plt} fixture exposes the Matplotlib \cite{hunter2007}
\texttt{pyplot} interface to a test function.
Test functions can then analyze and plot
the data generated in a simulation
(exposed from a simulator's \texttt{data} attribute)
to visualize the activity in a network.
Figures are saved in a specified or backend-specific directory.
The \texttt{logger} fixture exposes a logging interface
to save arbitrary text to a specified
or backend-specific directory;
these are useful for summary statistics
that can inspected manually.
For large amounts of data that
requires more analysis,
\texttt{analytics} fixture exposes an interface
to save arbitrary data (in the form of NumPy arrays)
to specified or backend-specific directories.

\paragraph{\texttt{analytics\_data}}
enables comparative testing between
two runs of the same test.
The two runs might represent two versions
of the same backend---one before a
speedup and one after---or two
entirely different backends.
The \texttt{analytics\_data} fixture
provides the results from those two
test runs to a single test function,
allowing for comparisons
between the two results saved
with the \texttt{analytics} fixture.
See Appendix~\ref{App:01} for a
concrete example of using the
\texttt{analytics\_data} fixture.

@@

\subsection{Metrics collected}
\label{Sec:Metrics}

\paragraph{Compliance}
is the number of tests that a particular backend
passes successfully, relative to the reference
backend, which must pass all tests.
While this metric can give an indication
of how many features a particular backend supports,
it does not take into account how commonly used
a particular feature is;
for this reason, a backend that has relatively
low compliance can still be useful
in many situations.

@@

It should be noted that
during a \texttt{run} call,
each backend incurs some overhead
to manage the simulation in addition
to actually moving the simulation forward.
The amount of overhead depends
on the backend; those on specific hardware platforms
(like the SpiNNaker and OpenCL backends)
are likely to incur more overhead
than those on general purpose platforms.
However, we include the overhead in
our speed benchmarking because
it is not possible to distinguish
overhead from actual run time
in a backend-agnostic way.

@@

Unlike the compliance metric,
accuracy and speed are only collected
for tests that define an accuracy metric
(i.e., those that use the \texttt{analytics} fixture).
In this study, we focus on four such tests
that address central functioning aspects
of a wide variety of large-scale briain models
(see \cite{eliasmith2003a,eliasmith2013}).

Each test has several parameters
that can be varied
in order to benchmark extreme cases
(see Appendix~\ref{App:01} for examples).
We test each model with parameters
typically used in large-scale models,
but make these parameterized models
available at \url{https://github.com/tbekolay/fnme2015}
for those who want to explore extreme cases.

@@

\paragraph{2. Two-dimensional product.}
In this model,
the scalar product is computed
from a two-dimensional ensemble
comprised of 100 LIF neurons.
Since the product is nonlinear,
we use a space-filling curve
(the Hilbert curve \cite{hilbert1891})
as the input signal to ensure that
we sample the entire two-dimensional space.
This model tests each backend's
ability to compute a nonlinear function,
albeit a low-dimensional one.
Despite the low dimensionality,
two-dimensional products are frequently used
in large-scale models;
it is the primary operation needed to support
symbol-like processing using neuron-like elements,
for example \cite{eliasmith2013}.

@@

\paragraph{4. Basal Ganglia sequence.}
In this model, brain structures
known collectively as the basal ganglia
are contructed from 4900 LIF neurons,
organized such that they iterate
through a repeating set of actions.
This model has been used to investigate
action selection and learning \cite{stewart2012a}
and the switching time between actions
has been closely mapped to human decision making
\cite{stewart2010}.
This benchmark tests the ability
to construct this model,
and evaluates the time needed
to transition between actions.

@@

\paragraph{1. Reference (\texttt{nengo}).}
The reference backend is designed to run quickly
on any general purpose computer
by using NumPy \cite{vanderwalt2011}
for fast vectorized computations.
It is included with the \texttt{nengo} frontend
as \texttt{nengo.Simulator}.
When appropriate, other backends will be benchmarked
relative to the reference backend,
which has the most features but does not target
specialized hardware.

@@

\paragraph{3. OpenCL (\texttt{nengo\_ocl}).}
The OpenCL backend uses the Open Computing Language
(OpenCL; \cite{stone2010})
to run Nengo models on many different computing devices,
including graphical processing units (GPUs),
and field-programmable gate arrays (FPGAs).
In contrast to the distilled backend,
it is designed to run Nengo models
as quickly as possible,
using fast general purpose computing devices
like GPUs and any optimizations available,
at the cost of code readability.

\paragraph{4. Brainstorm software (\texttt{nengo\_brainstorm}).}
The Brainstorm backend is a software implementation
of a new neuromorphic chip
based partly on Neurogrid \cite{benjamin2014}
currently in development
by the Brain in Silicon lab
at Stanford University\footnote{
  Some details available at \url{http://brainstorm.stanford.edu/projects/}.}.
The software backend does not aim for speed;
instead, it attempts to emulate the hardware
in order to test its applicability
to large-scale neural models.
If the emulated hardware can perform well,
then it follows that actual hardware
will also perform well,
but will be much faster.

\paragraph{5. SpiNNaker hardware (\texttt{nengo\_spinnaker}).}
The SpiNNaker backend \cite{mundyinpress}
targets the eponymous neuromorphic hardware
developed by \citeauthor{furber2014}.
In contrast to the Brainstorm backend,
this backend targets physical neuromorphic hardware,
and therefore is concerned
both with accuracy and speed.

\subsection{Benchmarking environment}

All benchmarks were run on a server
running Debian unstable.
The server was configured with
two Intel Xeon E5-2650 CPUs
clocked at 2.00GHz,
four Nvidia Tesla C2075 GPUs,
and 64 GB of RAM.
A single 48-chip SpiNN-5 board
was connected through
a 1 GB ethernet port on
the Supermicro X9DRG-QF motherboard.

The versions of software
used for benchmarking can be found in
Supplementary Table~1.

@@

Figure~\ref{fig:02} gives accuracy results
for the two-dimensional product model.
In this case, while all of the backends
perform well, the SpiNNaker backend
performs statistically worse than the
other four backends.
Its performance, however, is still well within
acceptable ranges for the two-dimensional product model.

Figure~\ref{fig:03} gives accuracy results
for the controlled oscillator model.
The reference and OpenCL backends perform well here,
as do the distilled and Brainstorm backends.
??? revisit with new results
%% The SpiNNaker backend performs well in most cases,
%% but some instances of the model
%% are signifless accurate
%% than the majority of instances.

Figure~\ref{fig:04} gives accuracy results
for the basal ganglia sequence model.
Again, all of the backends perform similarly,
except for the SpiNNaker backend.
All other backends have a transition time
around 43 milliseconds, which matches
the results from \citeauthor{stewart2010}
for simple actions.
The SpiNNaker backend has a transition time
around 55 milliseconds, which is slower
than the time reported by Stewart et al.
for simple actions, but faster than the
time reported for complex actions.
??? Run again with passthrough nodes removed
and make figure

@@

Figures~\ref{fig:05} and \ref{fig:06} show
the build and run speeds for all backends, respectively.
While build time is not critical to optimize,
it is worth noting that the reference backend
has consistently fast builds.
The OpenCL backend is also usually fast,
but can be slow for certain types of networks
(e.g., the chained communication channel model).
The distilled and Brainstorm backends---which
have a very similar build process---become
slow when dealing with moderately sized models
(e.g., the basal ganglia sequence model,
which contains 4900 neurons).
The SpiNNaker backend incurs an unavoidable cost
in setting up the hardware
(determining the placement of computational resources,
and generating routes to connect resouces),
though it builds the largest model faster than
the distilled and Brainstorm backends.

The run time results are of the utmost importance
when evaluating backends for different applications,
especially those that require real-time interaction,
such as robotics. In general, all backends
perform adequately for the three smaller models.
The OpenCL backend is significantly slower for small
models, like the product and controlled oscillator models.
SpiNNaker, on the other hand, is slower than other backends
for these models, because its goal
is to run in real-time even if it is possible to run faster.
The most important result is for the largest model,
the basal ganglia sequence.
In this model, the distilled and Brainstorm backends
performed very poorly
(though neither has speed as a primary goal).
The reference backend also did not perform well,
operating at nearly eight times slower than real-time.
The OpenCL backend performed well, operating at around
four times slower than real-time.
The best performance was seen by the SpiNNaker backend,
however, which operated at around 1.5 times real-time
on this moderately-sized model.
Since the SpiNNaker board always runs at real-time,
this means that around one-third of the time taken
for the \texttt{run} call is overhead.

Other backends also incur overhead,
especially when data is recorded
through probes.
We investigated the possibility
that the overhead involved in probes
dominated the actual run time for some backends.
However, we found that
removing probes from these models
made very little impact on
the total time taken in a \texttt{run} call
(see Supplementary Figure~1).

@@

The most important finding from these benchmarks
is that specialized hardware (including GPUs)
can result in significantly faster run times
for certain models,
with little to no cost in accuracy.
The conditions under which the OpenCL backend
performs slower than the reference backend
(e.g., the product and controlled oscillator models)
bear closer inspection to determine the source
of the slowdown.
Similarly, we should investigate
the model instances in which the SpiNNaker backend
differs from other backends
in the product model.

Above and beyond the benchmarking results,
however, we believe that the primary contribution
of this study is to provide evidence that Nengo
has a tested, stable, productive frontend
that can target multiple backends,
and therefore provides a platform
for benchmarking neuromorphic backends,
and other neural simulators.
While this is similar to the goal
of other projects, most notably PyNN \cite{davison2008},
Nengo is unique in focusing on
large-scale functional simulations,
rather than attempting to support
detailed single-neuron models
(though this capability is still possible in Nengo).
Topographica \cite{bednar2009} plays a similar role,
in that it can interact with multiple neural simulators
with a high-level API, but its API focuses specifically
on models of topographic maps
and other sensory pathways,
rather than focusing on function and dynamics.

@@

In summary, we have implemented
four benchmark models using the same
testing framework used for Nengo's test suite,
and gathered benchmark results
for five backends,
including an OpenCL backend that targets GPUs
and a backend that targets
SpiNNaker neuromorphic hardware.
All five backends were able to implement
the models accurately,
with the OpenCL and SpiNNaker backends
simulating the largest model much faster
than the reference backend.
We believe that the these benchmarks
can be easily expanded upon
to develop a suite of benchmarks
that can be run by any neuromorphic hardware
or neural simulator
with an associated Nengo backend.
Running these benchmarks in a common environment
and cleanly visualizing the results
could catalyze the development of
neural systems that implement
large-scale functional models
efficiently and accurately.

@@

\clearpage
\appendix

\section{Chained communication channel model}
\label{App:01}

\begin{lstlisting}
import numpy as np
import nengo
from nengo.utils.numpy import rmse

def test_cchannelchain(Simulator, analytics, plt, rng, seed):
    # Parameters that can be varied to investigate extreme cases
    dims = 2
    layers = 5
    n_neurons = 100
    synapse = nengo.Lowpass(0.01)

    with nengo.Network(seed=seed) as model:
        hypersphere = nengo.dists.UniformHypersphere()
        value = hypersphere.sample(dims, 1, rng=rng).ravel()
        stim = nengo.Node(value)

        ens = [nengo.Ensemble(n_neurons, dimensions=dims)
               for _ in range(layers)]

        nengo.Connection(stim, ens[0])
        for i in range(layers - 1):
            nengo.Connection(ens[i], ens[i+1], synapse=synapse)

        p_input = nengo.Probe(stim)
        p_outputs = [nengo.Probe(ens[i], synapse=synapse)
                     for i in range(layers)]

    sim = Simulator(model)
    sim.run(0.5)

    for p_output in p_outputs:
        plt.plot(sim.trange(), sim.data[p_output])
    plt.plot(sim.trange(), sim.data[p_input], color='k', linewidth=1)
    plt.ylabel('Decoded output')
    plt.xlabel('Time (s)')

    last = p_outputs[-1]
    decoding_rmse = rmse(value, sim.data[last][sim.trange() > 0.4]
    analytics.add_data('rmse', decoding_rmse)

def test_compare_cchannelchain(analytics_data, plt):
    rmses = [d['rmse'] for d in analytics_data]
    plt.bar(np.arange(len(rmses)), rmses, align='center')
    plt.ylabel("RMSE")
\end{lstlisting}

### 2015-07-01 paper/supplementary.tex
Full edit by Trevor
Filled in most sections now (except stuff at the end).
Ready to rerun benchmarks and regenerate plots.

>>>>

\begin{table}[!ht]
\textbf{\refstepcounter{table}\label{Tab:01} Supplementary Table \arabic{table}.}{
  Software versions used for benchmarking.
  The PyPI name is the unique identifier used for installing the Python package
  through the Python Package Index (PyPI). Packages can be found by
  visiting \url{https://pypi.python.org/pypi/<PyPI name>}.}

\processtable{ }
{\begin{tabular}{lcc}\toprule
 Package & PyPI name & Version number\\\midrule
 Nengo & \texttt{nengo} & Development version, commit \texttt{7d2d24145} \\
 Distilled backend & \texttt{nengo\_distilled} & 0.1.0 \\
 OpenCL backend & \texttt{nengo\_ocl} & 0.1.0 \\
 Brainstorm backend & Not available on PyPI & Development version, commit \texttt{ac3cfa708} \\
 SpiNNaker backend & \texttt{nengo\_spinnaker} & 0.2.3 \\
 NumPy & \texttt{numpy} & 1.8.1 \\\botrule
\end{tabular}}{}
\end{table}
